---
title: "MIT critical care"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
install.packages("RPostgreSQL")
install.packages("DBI")
library(DBI)
library(RPostgreSQL)
install.packages("reshape2")
library(reshape2)

install.packages("arules")
install.packages("arulesViz")
library(arules)
library(arulesViz)

install.packages("Hmisc")
library(Hmisc)
```

```{r}
# Load configuration settings
dbdriver <- 'PostgreSQL'
host  <- '18.194.208.218'
port  <- '5432'
user  <- 'user4'
password <- '2ooMXnc73d'
dbname <- 'mimic'
schema <- 'mimiciii'

# Connect to the database using the configuration settings
con <- dbConnect(dbDriver(dbdriver), dbname = dbname, host = host, port = port, 
                 user = user, password = password)

# Set the default schema
dbExecute(con, paste("SET search_path TO ", schema, sep=" "))

```



```{r}
#example query
sql_test_query <- "SELECT i.subject_id, i.hadm_id, i.los FROM icustays i;"
data <- dbGetQuery(con, sql_test_query)
```

SQL query for sepsis patients in MIMICiii dataset
-- ICD-9 codes for Angus criteria of sepsis

-- Angus et al, 2001. Epidemiology of severe sepsis in the United States
-- http://www.ncbi.nlm.nih.gov/pubmed/11445675

-- Case selection and definitions
-- To identify cases with severe sepsis, we selected all acute care
-- hospitalizations with ICD-9-CM codes for both:
-- (a) a bacterial or fungal infectious process AND
-- (b) a diagnosis of acute organ dysfunction (Appendix 2).
```{r}
sql_sepsis_query <- "WITH infection_group AS 
(
	SELECT subject_id, hadm_id,
	CASE
		WHEN substring(icd9_code,1,3) IN ('001','002','003','004','005','008',
			   '009','010','011','012','013','014','015','016','017','018',
			   '020','021','022','023','024','025','026','027','030','031',
			   '032','033','034','035','036','037','038','039','040','041',
			   '090','091','092','093','094','095','096','097','098','100',
			   '101','102','103','104','110','111','112','114','115','116',
			   '117','118','320','322','324','325','420','421','451','461',
			   '462','463','464','465','481','482','485','486','494','510',
			   '513','540','541','542','566','567','590','597','601','614',
			   '615','616','681','682','683','686','730') THEN 1
		WHEN substring(icd9_code,1,4) IN ('5695','5720','5721','5750','5990','7110',
				'7907','9966','9985','9993') THEN 1
		WHEN substring(icd9_code,1,5) IN ('49121','56201','56203','56211','56213',
				'56983') THEN 1
		ELSE 0 END AS infection
	FROM diagnoses_icd
),

organ_diag_group as
(
	SELECT subject_id, hadm_id,
		CASE
		-- Acute Organ Dysfunction Diagnosis Codes
		WHEN substring(icd9_code,1,3) IN ('458','293','570','584') THEN 1
		WHEN substring(icd9_code,1,4) IN ('7855','3483','3481',
				'2874','2875','2869','2866','5734')  THEN 1
		ELSE 0 END AS organ_dysfunction,
		-- Explicit diagnosis of severe sepsis or septic shock
		CASE
		WHEN substring(icd9_code,1,5) IN ('99592','78552')  THEN 1
		ELSE 0 END AS explicit_sepsis
	FROM diagnoses_icd
),

organ_proc_group as
(
	SELECT subject_id, hadm_id,
		CASE
		WHEN substring(icd9_code,1,4) IN ('9670','9671','9672') THEN 1
		ELSE 0 END AS mech_vent
	FROM procedures_icd
),

aggregate as
(
	SELECT subject_id, hadm_id,
		CASE
			WHEN hadm_id in
					(SELECT DISTINCT hadm_id
					FROM infection_group
					WHERE infection = 1)
				THEN 1
			ELSE 0 END AS infection,
		CASE
			WHEN hadm_id in
					(SELECT DISTINCT hadm_id
					FROM organ_diag_group
					WHERE explicit_sepsis = 1)
				THEN 1
			ELSE 0 END AS explicit_sepsis,
		CASE
			WHEN hadm_id in
					(SELECT DISTINCT hadm_id
					FROM organ_diag_group
					WHERE organ_dysfunction = 1)
				THEN 1
			ELSE 0 END AS organ_dysfunction,
		CASE
		WHEN hadm_id in
				(SELECT DISTINCT hadm_id
				FROM organ_proc_group
				WHERE mech_vent = 1)
			THEN 1
		ELSE 0 END AS mech_vent
	FROM admissions
)

SELECT subject_id, hadm_id, infection,
   explicit_sepsis, organ_dysfunction, mech_vent,
CASE
	WHEN explicit_sepsis = 1 THEN 1
	WHEN infection = 1 AND organ_dysfunction = 1 THEN 1
	WHEN infection = 1 AND mech_vent = 1 THEN 1
	ELSE 0 END
AS angus
FROM aggregate;"


df <- tbl_df(dbGetQuery(con, sql_sepsis_query))

sepsis <- df %>% 
  filter(explicit_sepsis == 1)
write.csv(sepsis, "~/IE-BIGDATA/Hackathons/MIT critical care datathon/sepsis.csv", fileEncoding = 'utf-8')
glimpse(sepsis)
```

```{r}
septicshock_query <- "WITH t1 AS (
SELECT 
  diagnoses_icd.subject_id,
  diagnoses_icd.icd9_code, 
  diagnoses_icd.hadm_id
FROM 
  mimiciii.d_icd_diagnoses, 
  mimiciii.diagnoses_icd
WHERE diagnoses_icd.icd9_code = '78552'
GROUP BY diagnoses_icd.subject_id, diagnoses_icd.icd9_code, diagnoses_icd.hadm_id
limit 50000
)

SELECT t1.subject_id, t1.hadm_id, count(*) as count
FROM t1
GROUP BY t1.subject_id, t1.hadm_id
HAVING count(*) = 1
;"

septickschock <- tbl_df(dbGetQuery(con, septicshock_query))

septicshock <- septickschock %>%
  select(-count)
```

```{r}
whitecells_query <- "WITH t1 AS (
SELECT 
  diagnoses_icd.subject_id,
  diagnoses_icd.icd9_code, 
  diagnoses_icd.hadm_id
FROM 
  mimiciii.d_icd_diagnoses, 
  mimiciii.diagnoses_icd
WHERE diagnoses_icd.icd9_code IN ('28800', '28801', '28802', '28803', '28804', '28809', '28953', '7767')
GROUP BY diagnoses_icd.subject_id, diagnoses_icd.icd9_code, diagnoses_icd.hadm_id
limit 50000
)

SELECT t1.subject_id, t1.hadm_id, count(*) as count
FROM t1
GROUP BY t1.subject_id, t1.hadm_id
HAVING count(*) = 1
;"

whiteblood <- tbl_df(dbGetQuery(con, whitecells_query))

whiteblood <- whiteblood %>%
  select(-count)
```

```{r}
dischargedeath_query <- "SELECT row_id, subject_id, hadm_id FROM mimiciii.admissions WHERE discharge_location = 'DEAD/EXPIRED';"

dischargedeath <- tbl_df(dbGetQuery(con, dischargedeath_query))
```


```{r}
#oasis score dataframe
oasisdf <- read_csv("OASIStabel.csv")
```

```{r}

#details icustay
icudetails <- read_csv("ICUdetails.csv")
#filter on age > 16 and 1st admission in ICU and Hospital
icudetails2 <- icudetails %>%
  filter(first_icu_stay=="Y") %>%
  filter(first_hosp_stay == "Y")
```

```{r}
#antibiotics info
antibiotics <- read_csv("ANTIBIOTICS.csv")

antibiotics <- antibiotics %>% select(subject_id, hadm_id, icustay_id, drug_formatted)

antibiotics <- dcast(antibiotics, subject_id + hadm_id + icustay_id ~ drug_formatted)
```

```{r}
#SOFA var
sofa <- read_csv("SOFA.csv")
```

Merging dataframes
```{r}
df1 <- merge(x=sepsis, y=oasisdf, by=c('subject_id','hadm_id') )

df2 <- merge(x=df1, y=icudetails2, by=c('subject_id','hadm_id', 'icustay_id') )



#fixed the duplicate issue by using semi_join. - 1510 records
df3 <- merge(x=df2, y=antibiotics, by=c('subject_id', 'hadm_id', 'icustay_id'))

fulldf <- df3 %>%
  select(-explicit_sepsis, -age.y, -preiculos_score, -age_score, -gcs_score, -heartrate_score, -meanbp_score, -resprate_score, -temp_score, -urineoutput_score, -mechvent_score, -electivesurgery_score, -ethnicity_1, -hospital_expire_flag.y, -first_hosp_stay, -first_icu_stay, -intime_1, -outtime_1, -angus, -icustay_age_group, -infection, -dob) %>% 
  rename(age=age.x) %>%
  rename(hospital_expire_flag = hospital_expire_flag.x)



#combine septicshock with dataframe
tempdf <- fulldf %>% 
  select(subject_id, hadm_id)
tempdf2 <- intersect(tempdf, septicshock) %>%
  mutate(septicshock = 1)
tempdf3 <- setdiff(tempdf, septicshock) %>%
  mutate(septicshock = 0)

tempdf4 <- bind_rows(tempdf2, tempdf3)
fulldf <- left_join(fulldf, tempdf4, by=c('subject_id', 'hadm_id'))

#combine whiteblood cells with datafram
tempdf4 <- intersect(tempdf, whiteblood) %>%
  mutate(whiteblood = 1)
tempdf5 <- setdiff(tempdf, whiteblood) %>%
  mutate(whiteblood = 0)

tempdf6 <- bind_rows(tempdf4, tempdf5)
fulldf <- left_join(fulldf, tempdf6, by=c('subject_id', 'hadm_id'))
fulldf <- fulldf %>%
  filter(whiteblood == 0) %>%
  select(-whiteblood)

fulldf <- fulldf %>%
  select(-dod)

fulldf <- fulldf %>%
  mutate(mortality = ifelse(is.na(deathtime), 0, 1))

sofa <- sofa %>%
  select(subject_id, hadm_id, icustay_id, sofa)

fulldf <- left_join(fulldf, sofa, by=c('subject_id', 'hadm_id', 'icustay_id'))


fulldf <- fulldf %>%
  select(-admittime, -dischtime, -deathtime, -icustay_seq, -intime, -outtime, -hospstay_seq)

fulldf <- fulldf %>%
  filter(`cefepime desensitization` == 0) %>%
  select(-`neo*iv*piperacillin-tazobactam`, -`cefepime desensitization`) %>%
  mutate(cefepime = ifelse(cefepime == 0, 0, 1)) %>%
  mutate(pipetazo = ifelse(`piperacillin-tazobactam` == 0, 0, 1)) 


fulldf <- fulldf %>%
  select(-`cefepime desensitization`)


write.csv(fulldf, "~/IE-BIGDATA/Hackathons/MIT critical care datathon/fulldf.csv", fileEncoding = 'utf-8')
```


cleaning data
```{r}
cleaneddf <- fulldf

cleaneddf$X <- NULL
cleaneddf$preiculos <- NULL
cleaneddf$hospital_expire_flag <- NULL
cleaneddf$icustay_expire_flag <- NULL
cleaneddf$mechvent <- NULL
cleaneddf$ethnicity <- NULL
cleaneddf$oasis <- scale(cleaneddf$oasis, center = FALSE)#o log?
cleaneddf$age <- scale(cleaneddf$age, center = FALSE)
cleaneddf$gcs <- scale(cleaneddf$gcs, center = FALSE)
cleaneddf$heartrate <- scale(cleaneddf$heartrate, center = FALSE)
cleaneddf$meanbp <- scale(cleaneddf$meanbp, center = FALSE)
cleaneddf$resprate <- scale(cleaneddf$resprate, center = FALSE)
cleaneddf$temp <- scale(cleaneddf$temp, center = FALSE)
cleaneddf$urineoutput <- scale(cleaneddf$urineoutput, center = FALSE)
cleaneddf$los_hospital <- scale(cleaneddf$los_hospital, center = FALSE)
cleaneddf$los_icu <- scale(cleaneddf$los_icu, center = FALSE)
cleaneddf$sofa <- scale(cleaneddf$sofa, center = FALSE)
cleaneddf$organ_dysfunction <- as.character(cleaneddf$organ_dysfunction)
cleaneddf[cleaneddf$organ_dysfunction=='0',]$organ_dysfunction <- 'N'
cleaneddf[cleaneddf$organ_dysfunction=='1',]$organ_dysfunction <- 'Y'
cleaneddf$mech_vent <- as.character(cleaneddf$mech_vent)
cleaneddf[cleaneddf$mech_vent=='0',]$mech_vent <- 'N'
cleaneddf[cleaneddf$mech_vent=='1',]$mech_vent <- 'Y'
cleaneddf$electivesurgery <- as.character(cleaneddf$electivesurgery)
cleaneddf[cleaneddf$electivesurgery=='0',]$electivesurgery <- 'N'
cleaneddf[cleaneddf$electivesurgery=='1',]$electivesurgery <- 'Y'
cleaneddf$cefepime <- as.character(cleaneddf$cefepime)
cleaneddf[cleaneddf$cefepime=='0',]$cefepime <- 'N'
cleaneddf[cleaneddf$cefepime=='1',]$cefepime <- 'Y'
cleaneddf$pipetazo <- as.character(cleaneddf$pipetazo)
cleaneddf[cleaneddf$pipetazo=='0',]$pipetazo <- 'N'
cleaneddf[cleaneddf$pipetazo=='1',]$pipetazo <- 'Y'
cleaneddf$septicshock <- as.character(cleaneddf$septicshock)
cleaneddf[cleaneddf$septicshock=='0',]$septicshock <- 'N'
cleaneddf[cleaneddf$septicshock=='1',]$septicshock <- 'Y'
cleaneddf$mortality <- as.character(cleaneddf$mortality)
cleaneddf[cleaneddf$mortality=='0',]$mortality <- 'ALIVE'
cleaneddf[cleaneddf$mortality=='1',]$mortality <- 'DEAD'


#make factors /logicals

cleaneddf$organ_dysfunction <- as.factor(cleaneddf$organ_dysfunction)

cleaneddf$mech_vent <- as.factor(cleaneddf$mech_vent)

cleaneddf$electivesurgery <- as.factor(cleaneddf$electivesurgery)

cleaneddf$cefepime <- as.factor(cleaneddf$cefepime)

cleaneddf$pipetazo <- as.factor(cleaneddf$pipetazo)

cleaneddf$septicshock <- as.factor(cleaneddf$septicshock)

cleaneddf$mortality <- as.factor(cleaneddf$mortality)

cleaneddf$gender <- as.factor(cleaneddf$gender)

cleaneddf$admission_type <- as.factor(cleaneddf$admission_type)



```



visual analysis
```{r}
ggplot(data=fulldf, aes(y=los_icu, x=septicshock)) +
  geom_point() +
  geom_smooth(method=lm)

```


Apriori rule finding
```{r}

apriori <- cleaneddf %>%
  select(-subject_id, -hadm_id, -icustay_id)

apriori <- apriori %>%
  filter(is.na(heartrate) == F) %>%
  filter(is.na(temp)==F) %>%
  filter(is.na(urineoutput) == F)

q1 <- function(x){
  quantile(x, 0.2)
}
q2 <- function(x){
  quantile(x, 0.4)
}
q3 <- function(x){
  quantile(x, 0.6)
}
q4 <- function(x){
  quantile(x, 0.8)
}

apriori <- apriori %>%
  mutate(oasis = ifelse(apriori$oasis < q1(apriori$oasis), 1, 
                        ifelse(apriori$oasis < q2(apriori$oasis),2,
                               ifelse(apriori$oasis < q3(apriori$oasis), 3, 
                                      ifelse(apriori$oasis < q4(apriori$oasis), 4, 5)))))
apriori <- apriori %>%
  mutate(age = ifelse(apriori$age < q1(apriori$age), 1, 
                        ifelse(apriori$age < q2(apriori$age),2,
                               ifelse(apriori$age < q3(apriori$age), 3, 
                                      ifelse(apriori$age < q4(apriori$age), 4, 5))))) %>%
  mutate(heartrate = ifelse(apriori$heartrate < q1(apriori$heartrate), 1, 
                        ifelse(apriori$heartrate < q2(apriori$heartrate),2,
                               ifelse(apriori$heartrate < q3(apriori$heartrate), 3, 
                                      ifelse(apriori$heartrate < q4(apriori$heartrate), 4, 5))))) %>%
  mutate(meanbp = ifelse(apriori$meanbp < q1(apriori$meanbp), 1, 
                        ifelse(apriori$meanbp < q2(apriori$meanbp),2,
                               ifelse(apriori$meanbp < q3(apriori$meanbp), 3, 
                                      ifelse(apriori$meanbp < q4(apriori$meanbp), 4, 5))))) %>%
  mutate(resprate = ifelse(apriori$resprate < q1(apriori$resprate), 1, 
                        ifelse(apriori$resprate < q2(apriori$resprate),2,
                               ifelse(apriori$resprate < q3(apriori$resprate), 3, 
                                      ifelse(apriori$resprate < q4(apriori$resprate), 4, 5))))) %>%
  mutate(temp = ifelse(apriori$temp < q1(apriori$temp), 1, 
                        ifelse(apriori$temp < q2(apriori$temp),2,
                               ifelse(apriori$temp < q3(apriori$temp), 3, 
                                      ifelse(apriori$temp < q4(apriori$temp), 4, 5))))) %>%
  mutate(urineoutput = ifelse(apriori$urineoutput < q1(apriori$urineoutput), 1, 
                        ifelse(apriori$urineoutput < q2(apriori$urineoutput),2,
                               ifelse(apriori$urineoutput < q3(apriori$urineoutput), 3, 
                                      ifelse(apriori$urineoutput < q4(apriori$urineoutput), 4, 5))))) %>%
  mutate(los_hospital = ifelse(apriori$los_hospital < q1(apriori$los_hospital), 1, 
                        ifelse(apriori$los_hospital < q2(apriori$los_hospital),2,
                               ifelse(apriori$los_hospital < q3(apriori$los_hospital), 3, 
                                      ifelse(apriori$los_hospital < q4(apriori$los_hospital), 4, 5))))) %>%
  mutate(los_icu = ifelse(apriori$los_icu < q1(apriori$los_icu), 1, 
                        ifelse(apriori$los_icu < q2(apriori$los_icu),2,
                               ifelse(apriori$los_icu < q3(apriori$los_icu), 3, 
                                      ifelse(apriori$los_icu < q4(apriori$los_icu), 4, 5))))) %>%
  mutate(sofa  = ifelse(apriori$sofa  < q1(apriori$sofa ), 1, 
                        ifelse(apriori$sofa  < q2(apriori$sofa ),2,
                               ifelse(apriori$sofa  < q3(apriori$sofa ), 3, 
                                      ifelse(apriori$sofa  < q4(apriori$sofa ), 4, 5)))))
 
apriori2 <- apriori %>%
  select(-oasis_prob, -gcs) 


apriori2$oasis <- as.factor(apriori2$oasis)
apriori2$age <- as.factor(apriori2$age)
apriori2$heartrate <- as.factor(apriori2$heartrate)
apriori2$meanbp <- as.factor(apriori2$meanbp)
apriori2$resprate <- as.factor(apriori2$resprate)
apriori2$temp <- as.factor(apriori2$temp)
apriori2$los_hospital <- as.factor(apriori2$los_hospital)
apriori2$los_icu <- as.factor(apriori2$los_icu)
apriori2$sofa <- as.factor(apriori2$sofa)
apriori2$urineoutput <- as.factor(apriori2$urineoutput)





rules <- apriori(apriori2, parameter=list(support =0.01, confidence =0.5, minlen=2, maxlen=5), appearance = list(rhs=c("mortality=DEAD", "mortality=ALIVE"), lhs=c('pipetazo=Y', 'cefepime=Y')))

rulesdf <- as(rules, "data.frame")
inspect(head(sort(rules, by="lift"),20))

plot(rules)

plot(rules, measure=c("support","lift"), shading="confidence")

subrules <-  head(sort(rules, by="lift"), 30)
plot(subrules, method="graph")
oneRule = sample(rules, 1)
inspect(oneRule)

inspect(head(subset(rules, lhs %pin% "mortality"), 10))


write.csv(apriori2, "~/IE-BIGDATA/Hackathons/MIT critical care datathon/apriori2.csv", fileEncoding = 'utf-8')

```

```{r}
#t.tests and regression


```


